```{r}
library(dplyr)
library(sf)
library(pracma)
library(tmap)

source("weave-grids.R")
```

# Get a grid generator and apply it 

```{r}
gg2 <- grid_generator(S = 100)
gg2(c(1, 1))

gg3 <- grid_generator(n_axes = 3)
gg3(c(1, 1, 1))
```

# Cell basis polygon
Some example grid cell polygons

```{r}
get_grid_cell_polygon(n_sides = 3) %>% plot(xlim = c(-0.6, 0.6), ylim = c(-0.55, 0.7))
get_grid_cell_polygon(n_sides = 4) %>% plot(add = TRUE)
get_grid_cell_polygon(n_sides = 5) %>% plot(add = TRUE, border = "grey") # because we can
get_grid_cell_polygon(n_sides = 6) %>% plot(add = TRUE)
```

# Slices that cut a cell strand along its 'length'

```{r}
# get a cell polygon
cell <- get_grid_cell_polygon(n = 3) %>% st_sfc()
# bounding box to determine any offset between its centroid (at 0,0) and its vertical centre
bb <- st_bbox(cell)
offset <- c(bb$xmin + bb$xmax, bb$ymin + bb$ymax) / 2
# get slices
slices <- get_grid_cell_slices(L = bb$xmax - bb$xmin, W = 0.5, 
                               n_slices = 2, offset = offset)
cell %>% plot()
slices %>% plot(add = TRUE)
```



# Get strand polygons
There are `get_cell_strands()` and `get_all_cell_strands()` functions.

```{r}
cell <- get_grid_cell_polygon(n_sides = 3) %>% st_sfc()
strands <- get_all_cell_strands(n = 3, width = 0.25, 
                                orientations = c(0,120,240), n_slices = 1:3)
cell %>% plot()
strands %>% plot(add = TRUE, col = rgb(0,0,0,0.2))
```

```{r}
cell <- get_grid_cell_polygon(n = 3) %>% rotate_shape(180) %>% st_sfc()
strands <- get_cell_strands(n = 3, width = 0.5, orientation = 180)
cell %>% plot()
strands %>% plot(add = TRUE, col = "grey")
```
# Using `weave-grids.R` to make weave drawings
The approach is to make a grid generator, then a set of coordinates, then apply positional offsets from the coordinates returned by the generator to the base cell strands.

For a triangular grid only coordinate combinations that sum to two consecutive values exist, and the parity of the sum determines the facing (up or down) of the triangular cell.

```{r}
trigrid <- grid_generator(n_axes = 3)
coords3 <- expand.grid(a = -3:3, b = -1:1, c = -1:1)
coords3 <- coords3 %>%
  mutate(parity = a + b + c) %>% 
  filter(parity %in% 0:1) # in a triaxial grid only these coordinate combos exist Nagy 2003
centres3 <- apply(coords3[, letters[1:3]], 1, trigrid) %>% t()
```

Now use the grid centres to generate the strands.

```{r}
trigridcells <- list()
permutations <- pracma::perms(1:3) # all possible orderings
orientations <- c()   # we save these for colouring
strands <- c()        
for (i in 1:nrow(centres3)) {
  # permutation would eventually be determined by some function based on user input
  permutation <- permutations[sample(1:6, 1), ] %>% c()
  o <- ifelse(coords3$parity[i] == 1, 0, 180) +          # which way up is it
    c(0, 120, 240)[permutation]                          # rotation depending on the permutation
  n <- 1
  strands <- c(strands, rep(letters[1:3][permutation], n) %>%
                 matrix(n, 3, byrow = TRUE) %>% c())   # record the strand
  next_strands <- get_all_cell_strands(n = 3, width = 0.333, n_slices = rep(n, length(o)),
                                       orientations = o) + centres3[i, ]
  for (p in next_strands) {
    trigridcells <- append(trigridcells, list(p))
  }
}
trigridcells <- trigridcells %>%
  st_sfc()
```

Now make into an `sf` and plot

```{r}
tgc <- data.frame(strand = strands) %>%
  st_sf(geometry = trigridcells)
tm_shape(tgc) +
  tm_polygons(col = "strand", palette = "Set1", border.col = "grey") + 
  tm_layout(main.title = "Note: not a weavable pattern")
```

# Biaxial grid
This is similar, but simpler...

```{r}
bigrid <- grid_generator(S = 100)
ab <- expand.grid(-5:5, -5:5)
centres2 <- apply(ab, 1, bigrid) %>% t()
bigridcells <- list()
for (i in 1:nrow(centres2)) {
  if (i %% 2 == 0) {
    o <- c(0, 90)
  } else {
    o <- c(90, 0)
  }
  next_strands <- get_all_cell_strands(S = 100, width = 0.75, 
                                      orientations = o, n_slices = c(1, 1)) + centres2[i, ]
  for(p in next_strands) {
    bigridcells <- append(bigridcells, list(p))
  }
}
bigridcells <- bigridcells %>%
  st_sfc()
tm_shape(bigridcells) +
  tm_polygons()
```

