# Combining orderings from biaxial weaves
This function takes a list of orderings on the values 1:n, and combines them into a single consistent ordering. If a value is not present in any of the members of the orderings list it will not appear in the output. If all the orderings are empty, `NULL` is returned.

```{r}
# combines a set of orderings on the values
# the orderings are a list of vectors (which may be empty)
# for example list(c(1, 2), c(2, 3), c(1, 3))
combine_orderings <- function(orderings, values = 1:3, verbose = FALSE) {
  # assemble a matrix of the positions of entries
  # in each ordering among the values
  ranks <- matrix(0, length(orderings), length(values))
  for (i in seq_along(orderings)) {
    ranks[i, ] <- match(values, orderings[[i]])
  }
  # replace any missing matches with a high score
  ranks[which(is.na(ranks))] <- 100
  # sum the ranks of each value
  scores <- colSums(ranks)
  max_score <- length(orderings) * 100
  number_present <- sum(scores < max_score)
  if (number_present == 0) {
    result <- NULL 
  } else {
    result <- values[order(scores)[1:number_present]]
  }
  if (verbose) {
    return(list(result = result, ranks = ranks, scores = scores)) 
  } else {
    return(result)
  }
}
```

## Examples
### All three strands present
Note that not all combinations of inputs in `orderings` should arise

```{r}
# 3rd entry MUST be 1,3 because cyclic is impossible
combine_orderings(list(1:2, 2:3, c(1,3)), verbose = TRUE) 
```

Some more
```{r}
combine_orderings(list(1:2, 3:2, c(1,3))) 
combine_orderings(list(2:1, 2:3, c(1,3))) 
combine_orderings(list(2:1, 2:3, c(3,1))) 
combine_orderings(list(2:1, 3:2, c(3,1))) 
```

### One thread missing
Here the missing thread will be missing twice, and two of the inputs will contain only the other thread. Note that function can accept single values where only one strand is present in an input.

```{r}
combine_orderings(list(1, 3, c(1,3))) 
combine_orderings(list(1:2, 2, 1)) 
combine_orderings(list(2, 2:3, 3)) 
```

### Two threads missing
Here, the thread that is present should appear twice, once in each constituent weave

```{r}
combine_orderings(list(1, NULL, 1)) 
```

### All threads missing
```{r}
combine_orderings(list(NULL, NULL, NULL)) 
```

### Biaxial case
This should also work... although we wouldn't necessarily want to use it.

```{r}
combine_orderings(list(2:1)) 
combine_orderings(list(1))
combine_orderings(list(NULL))
```

## More threads?
Note sure if this will also work, but it _might_! E.g., 

```{r}
combine_orderings(list(1:2, 3:2, 3:4, c(1,4)), 1:4, verbose = TRUE)
```
