---
output:
  pdf_document: default
  html_document: default
---
```{r results = FALSE, message = FALSE, warning = FALSE}
source("weaving-space-utils.R")
source("biaxial-weave-units.R")
source("triaxial-weave-units.R")
source("weave-map.R")

library(sf)        # vector spatial data
library(tmap)      # thematic maps
library(dplyr)     # data wrangling
```

# Plain weave with gaps

```{r}
m <- make_twill_matrix(1) %>%
  repmat(2) %>%
  augment_with_values(2, 0)
unit <- get_biaxial_weave_unit(type = "this", tie_up = m, strands = "abcd--|efgh--")
unit$primitive %>% plot(border = NA)
```

Make the gap bigger - any time you add more rows/columns of 0s with `augment_with_values` you need to add dashes to the `strands` parameter.

```{r}
m <- make_twill_matrix(1) %>%
  repmat(2) %>%
  augment_with_values(3, 0)
unit <- get_biaxial_weave_unit(type = "this", tie_up = m, strands = "abcd---|efgh---")
unit$primitive %>% plot(border = NA)
```

# 2 over 2 under basket weave with gaps

```{r}
m <- make_basket_matrix(2) %>% 
  augment_with_values(2, 0)
unit <- get_biaxial_weave_unit(type = "this", tie_up = m, strands = "abcd--|efgh--")
unit$primitive %>% plot(border = NA)
```

This can be augmented similarly

```{r}
m <- make_basket_matrix(2) %>% 
  augment_with_values(3, 0)
unit <- get_biaxial_weave_unit(type = "this", tie_up = m, strands = "abcd---|efgh---")
unit$primitive %>% plot(border = NA)
```

If you wanted to have spaces 4x4 then since the 2 over 2 under basket fits that exactly, you don't need to augment the matrix, and can just add the missing strands:

```{r}
m <- make_basket_matrix(2)
unit <- get_biaxial_weave_unit(type = "this", tie_up = m, strands = "abcd----|efgh----")
unit$primitive %>% plot(border = NA)
```
